Assessing the Replicability of Psychological Science Through Pedagogy

### 2016-08-03 paper/paper.Rmd
some minor paper/bib updates

---
title: "Assessing the Replicability of Psychological Science Through Pedagogy"
short-title: "Replication Through Pedagogy"
output: kmr::apa_manuscript
csl: apa6.csl
bibliography: replication254.bib

@@

abstract: 
    "Replications are important to science, but who will do them? One proposal is that students can conduct direct replications as part of their training; but to date there is limited evidence that this proposal can result in high-quality research outputs. We report a series of 11 pre-registered, direct replications of findings from the 2015 volume of Psychological Science, all conducted as part of a graduate-level course. The modal outcome was a partial replication, with most yielding weaker evidence than the original study. This work provides an estimate of the probability that a motivated graduate student can reproduce a previously published finding of interest on a first attempt. We describe the workflow and pedagogical methods that were used in the class (providing access to all materials) and discuss challenges for adoption of this model more broadly."

@@

Reproducibility is a core value for empirical research, but there is increasing concern throughout psychology that weak statistical standards and methodological practices have led to low empirical rates of replication [@osc2015]. Under the current incentive structure for science, direct replication is not typically valued, however. One potential solution to this problem is to make direct replication an explicit part of pedagogy: that is, to teach students about experimental methods by asking them to run replication studies [@frank2012;@grahe2012]. Despite enthusiasm for this idea, however [@lebel2015;@everett2015;@standing2016;@king2016], there is limited data -- beyond anecdotal reports or individual projects [@phillips2015;@lakens2013] -- to support its efficacy.

@@

It is a challenging proposition to estimate the replicability of an entire literature [cf. @osc2015]. Making such an estimate requires determining a sampling strategy, a notion of what constitutes a direct replication, what constitutes a successful replication, among many other difficult decisions [@gilbert2016;@anderson2016]. In the current work, we focus on a quantity that is easier to estimate: the probability that a graduate student can choose an article of interest in *Psychological Science* and -- in a single attempt, within constraints of budget, expertise, and effort -- reproduce the basic pattern of findings with sufficient fidelity to be able to build on it in future work. We believe that this notion of whether a piece of research supports cumulative progress is an important construct. Whether another independent scientist can build on a result is perhaps as important as -- and in many cases easier to measure -- whether the work is reproducible in some more abstract sense.

@@

d <- left_join(project_plan, project_data_summary, by = c("project_key"))
```

@@

All study participants were recruited on Amazon Mechanical Turk; individual samples are given in Table \ref{tab:summary}. Note that of the `r nrow(ppt)` studies included in the final sample, `r sum(project_plan$original_on_turk=="Yes")` (`r 100 * round(mean(project_plan$original_on_turk=="Yes"), digits=2)`%) were originally conducted on Mechanical Turk. All experiments were approved by the Stanford University institutional review board under protocol \#23274, "Reproducibility of psychological science and instruction."

@@

Of the `r nrow(ppt)` studies included in the final sample, `r 100 * round(mean(project_plan$open_materials == "Yes" | project_plan$open_materials == "Some"), digits=2)`% of studies had either some or all materials openly available. For `r sum(project_plan$author_contact != "")` studies, however, students contacted the authors to request either materials or clarification of methods (using a template email that was customized by the students and reviewed by the instructors). Responses were received in all but one case, usually within a matter of days.

@@

We eventually used a hybrid, case-by-case selection criterion that attempted to maximize the success of the project while staying within the constraints of our budget.
^[We initially were allocated \$1,500 but one student contributed personal research funds, leading to a total cost of $\approx$ \$1,700 for all studies. We attempted to set payment for our studies at approximately \$6/hour based on timing estimated from pilot A sessions.] 
In 
`r sum(project_plan$power_standard == "80% power")`
cases, we powered the replication attempt to 80\% post-hoc power. In 
`r sum(project_plan$power_standard == "Original")` 
cases, we used the original sample size. In one case, the original experiment was apparently extremely over-powered in the post-hoc analysis, so we chose a smaller sample size that still would apparently yield high power.  In one case, we reduced the number of trials as part of the adaptation to mechanical turk, and in one case of a multi-step procedure where the first step consisted of stimulus generation, we decreased the N in the first stage of the procedure (details available in Supplemntal Materials).

@@

Details of individual replications are available in the supplementary materials. Students chose findings from a wide variety of domains for replication. One cluster of studies included investigations of memory and visual attention [e.g., @xu2015], including in applied contexts like remembering "deleted" files [@storm2015] or tracking transition between locations on smartphone apps [@liverence2015]. A second cluster came from studies of social perception, including judgments of faces and voices [@ko2015;@sofer2015], as well as studies of attributions of modesty, creativity, and expertise [@atir2015;@scopelliti2015;@proudfoot2015]. Other studies investigated increasing retirement savings by orienting participants to the future [@lewis2015], legacy priming [@zaval2015], and the effects of math anxiety on performance [@wang2015].

@@

Of the 15 students in the class, all completed projects (two with extensions), but one failed to find a project of interest in the sampling frame and three opted out of the broader project prior to data collection.^[Note that these students did not express doubts about the success of their projects prior to opting out, and  2/3 projects were judged by the instructor team to have been successful replications. Thus we have no evidence that these students opted out systematically due to their belief that their project would fail.] The remaining 11 students contributed code, data, registrations and materials to the final product. (In one case, a student failed to complete the pre-registration but nevertheless submitted a pre-specified analysis plan to the instructors for review).

The instructor team coded the fidelity of replications, weighing whether the materials, sample, and task parameters differed from the original studies. Coded on a scale of 1 -- 7 (with 1 being a loose replication with substantive deviations from the original, and 7 being essentially identical), the modal rating was `r round(mean(d$replication_fidelity_MCF), digits = 2)`. Several studies were essentially identical to the originals, but there was a group of others that included differences in population or number of trials that might have an effect on the results.

@@

```{r}
d$high_fidelity <- d$replication_fidelity_MCF > median(d$replication_fidelity_MCF)
# table(d$replicated_TA,d$replicated_student) # Make prettier
#mean(d$replicated_team)

@@

- key statistic in original

>>>>

---
title: "Assessing the Replicability of Psychological Science Through Pedagogy"
short-title: "Replication Through Pedagogy"
output: kmr::apa_manuscript

csl: apa6.csl
bibliography: replication254.bib

@@

abstract: 
    "Replications are important to science, but who will do them? One proposal is that students can conduct direct replications as part of their training. As a proof-of-concept for this idea, here we report a series of 11 pre-registered, direct replications of findings from the 2015 volume of Psychological Science, all conducted as part of a graduate-level course. This work provides an estimate of the probability that a motivated graduate student can reproduce a previously published finding of interest on a first attempt. Congruent with previous findings, replication studies typically yielded smaller effects than originals, though the modal outcome was some support for the original claim. We describe the workflow and pedagogical methods that were used in the class and discuss challenges for adoption of this pedagogical model and replication research more broadly."

@@

Replicability is a core value for empirical research and there is increasing concern throughout psychology that more independent replication is necessary [@osc2015;@wagenmakers2012]. Under the current incentive structure for science, direct replication is not typically valued, however. One potential solution to this problem is to make direct replication an explicit part of pedagogy: that is, to teach students about experimental methods by asking them to run replication studies [@frank2012;@grahe2012]. Despite enthusiasm for this idea, however [@lebel2015;@everett2015;@standing2016;@king2016], there is limited data -- beyond anecdotal reports or individual projects [@phillips2015;@lakens2013] -- to support its efficacy.

@@

It is a challenging proposition to estimate the replicability of an entire literature [cf. @osc2015]. Making such an estimate minimally requires determining a sampling strategy, a notion of what constitutes a direct replication, and a standard for what constitutes a successful replication, among many other difficult decisions [@gilbert2016;@anderson2016]. In the current work, we focus on a quantity that is easier to estimate: the probability that a graduate student can choose an article of interest in *Psychological Science* and -- in a single attempt, within constraints of budget, expertise, and effort -- reproduce the basic pattern of findings with sufficient fidelity to be able to build on it in future work. We believe that this notion of whether a piece of research supports cumulative progress is an important construct. Whether another independent scientist can build on a result is perhaps as important as -- and in many cases easier to measure -- whether the work is reproducible in some more abstract sense.

@@

d <- left_join(project_plan, project_data_summary, by = c("project_key")) %>%
  rowwise() %>%
  mutate(fidelity = (replication_fidelity_MCF + replication_fidelity_ES) / 2) %>%
  ungroup
```

@@

All study participants were recruited on Amazon Mechanical Turk; individual samples are given in Table \ref{tab:summary}.^[Note that of the `r nrow(ppt)` studies included in the final sample, `r sum(project_plan$original_on_turk=="Yes")` (`r 100 * round(mean(project_plan$original_on_turk=="Yes"), digits=2)`%) were originally conducted on Mechanical Turk.] All experiments were approved by the Stanford University institutional review board under protocol \#23274, "Reproducibility of psychological science and instruction."

@@

Of the `r nrow(ppt)` studies included in the final sample, `r 100 * round(mean(project_plan$open_materials == "Yes" | project_plan$open_materials == "Some"), digits=2)`% of studies had either some or all materials openly available. For `r sum(project_plan$author_contact != "")` studies, however, students contacted the authors to request either materials or clarification of methods (using a template email that was customized by the students and reviewed by the instructors). Responses were received in all but one case, all within a matter of days.

@@

We eventually used a hybrid, case-by-case selection criterion that attempted to maximize the success of the project while staying within the constraints of our budget.
^[We initially were allocated \$1,500 but one student contributed personal research funds, leading to a total cost of $\approx$ \$1,700 for all studies. We attempted to set payment for our studies at approximately \$6/hour based on timing estimated from pilot A sessions.] 
In 
`r sum(project_plan$power_standard == "80% power")`
cases, we powered the replication attempt to 80\% post-hoc power. In 
`r sum(project_plan$power_standard == "Original")` 
cases, we used the original sample size. In one case, the original experiment was apparently extremely over-powered in the post-hoc analysis, so we chose a smaller sample size that still would apparently yield high power.  In one case, we reduced the number of trials as part of the adaptation to mechanical turk, and in one case of a multi-step procedure where the first step consisted of stimulus generation, we decreased the N in the first stage of the procedure (details available in SOM).

@@

Students chose findings from a wide variety of domains for replication. One cluster of studies included investigations of memory and visual attention [e.g., @xu2015], including in applied contexts like remembering "deleted" files [@storm2015] or tracking transition between locations on smartphone apps [@liverence2015]. A second cluster came from studies of social perception, including judgments of faces and voices [@ko2015;@sofer2015], as well as studies of attributions of modesty, creativity, and expertise [@atir2015;@scopelliti2015;@proudfoot2015]. Other studies investigated increasing retirement savings by orienting participants to the future [@lewis2015], legacy priming [@zaval2015], and the effects of math anxiety on performance [@wang2015]. Details of individual replications are available in SOM.

@@

Of the 15 students in the class, all completed projects (two with extensions), but one failed to find a project of interest in the sampling frame and three opted out of the broader project prior to data collection.^[Note that these students did not express doubts about the success of their projects prior to opting out, and 2/3 projects were judged by the instructor team to have been successful replications. Thus we have no evidence that these students opted out systematically due to their belief that their project would fail.] The remaining 11 students contributed code, data, registrations and materials to the final product. (In one case, a student failed to complete the pre-registration procedure correctly but nevertheless submitted a pre-specified analysis plan to the instructors for review).

Each member of the instructor team coded the fidelity of replications, weighing whether the materials, sample, and task parameters differed from the original studies. Coded on a scale of 1 -- 7 (with 1 being a loose replication with substantive deviations from the original, and 7 being essentially identical), the mean rating was `r round(mean(d$fidelity), digits = 2)`. Several studies were essentially identical to the originals, but there was a group of others that included differences in population or number of trials that might have an effect on the results.

@@

```{r}
d$high_fidelity <- d$fidelity > median(d$fidelity)
# table(d$replicated_TA,d$replicated_student) # Make prettier
#mean(d$replicated_team)

@@

key statistic in original
